\begin{abstract}
Tax-benefit microsimulation models typically operate at the national level, using household survey weights calibrated to aggregate population targets. Subnational analysis---at the level of states, congressional districts, or local authorities---requires datasets that simultaneously satisfy geographic distributional constraints while preserving household-level detail. We present a method based on $L_0$ regularization that jointly optimizes survey weight magnitudes and sparsity to produce calibrated subnational microsimulation datasets.

Our approach builds on the Hard Concrete distribution \citep{louizos2018}, which induces exact sparsity by multiplying each household's weight by a learned stochastic gate that collapses to a deterministic zero or one at inference time. We parameterize each gate with a log-alpha and temperature parameter, and jointly optimize these alongside log-transformed weight magnitudes using a single loss function combining scale-invariant relative calibration error, an $L_0$ sparsity penalty on the expected count of active households, and a light $L_2$ regularizer on weight magnitudes.

The pipeline begins with the US Current Population Survey. Each household record is cloned multiple times and assigned to random census blocks drawn from a population-weighted distribution. Program participation indicators are re-randomized per geographic assignment using local take-up rates. Each clone is then run through \policyengine{}'s tax-benefit microsimulation engine to generate geography-specific outputs. The $L_0$ optimizer selects which household-geography combinations to retain, calibrating simultaneously against approximately 37,800 targets across three geographic levels. The sparsity penalty is configurable: a higher penalty produces a compact national dataset of approximately 50,000 records, while a lower penalty yields a larger dataset of approximately 3--4 million records covering all 436 congressional districts and 50 states individually. The method is implemented as the open-source \texttt{l0-python} PyTorch package.
\end{abstract}

\noindent\textbf{Keywords:} microsimulation, survey calibration, $L_0$ regularization, subnational analysis
