\section{Introduction}
\label{sec:introduction}

Microsimulation models estimate the effects of tax and benefit policies on households by applying program rules to individual-level microdata. Most operational models---including those maintained by the Congressional Budget Office \citep{cbo2018}, the Joint Committee on Taxation \citep{jct2023}, and the Tax Policy Center \citep{tpc2024}---operate at the national level. They calibrate household survey weights to aggregate administrative totals such as total income tax revenue, program enrollment counts, and demographic benchmarks, then use the reweighted dataset to simulate policy reforms.

Subnational policy analysis introduces a fundamentally different calibration challenge. Rather than matching a single set of national aggregates, the microdata must simultaneously reproduce distributional statistics at multiple geographic levels: congressional districts, states, and the nation as a whole. A dataset calibrated for the state of California must match California-specific IRS income totals, SNAP participation counts, Medicaid enrollment, and age distributions, while remaining consistent with national budget projections from the CBO and tax expenditure estimates from the JCT. Across 436 congressional districts and 50 states, this produces approximately 37,800 simultaneous calibration targets.

Existing calibration methods scale poorly to this setting. Iterative proportional fitting \citep[IPF;][]{deming1940, ireland1968} adjusts weights along one dimension at a time, cycling through marginal constraints until convergence. IPF handles cross-classified tables but does not naturally accommodate hierarchical geographic constraints---district targets must sum to state targets, which must sum to national targets---without ad hoc post-processing. Generalized regression (GREG) estimators \citep{deville1992, sarndal2007} solve a constrained optimization problem that minimizes distance from initial weights subject to exact calibration constraints. GREG produces a closed-form solution for moderate numbers of constraints but becomes computationally intractable and numerically unstable as the constraint count approaches the tens of thousands.

Spatial microsimulation methods take a different approach, constructing synthetic populations for small areas by combinatorial optimization \citep{williamson1998, huang2001}, simulated annealing \citep{harland2012}, or deterministic reweighting \citep{tanton2011, lovelace2016}. These methods typically operate at a single geographic level and require separate calibration runs for each area, making joint multi-level calibration difficult.

This paper presents a method that addresses these limitations by jointly optimizing weight magnitudes and sparsity in a single gradient-based framework. We adapt the Hard Concrete distribution \citep{louizos2018}, originally developed for neural network pruning, to the survey calibration setting. Each household-geography combination receives a continuous weight and a stochastic binary gate. The gate is parameterized by a learnable logit and trained via gradient descent to minimize a loss function that combines relative calibration error across all 37,800 targets with an $L_0$ penalty on the expected number of active records. At inference time, the stochastic gates collapse to deterministic zeros and ones, producing a sparse dataset in which most household-geography combinations are dropped while the retained records carry calibrated positive weights.

The approach builds on \citet{woodruff2024}, who developed a two-stage methodology for constructing enhanced national microsimulation datasets from the Current Population Survey (CPS) and the IRS Public Use File (PUF). Their method uses quantile regression forests (QRF) to impute 72 tax variables from the PUF onto CPS records, then applies dropout-regularized gradient descent to reweight the combined dataset against approximately 7,000 national targets. The present paper extends this framework from a single national dataset to subnational coverage by introducing three new components: (a) a clone-and-assign procedure that replicates each CPS household across multiple geographic locations, (b) $L_0$ Hard Concrete gates that replace dropout regularization and enable exact sparsity, and (c) a hierarchical uprating scheme that reconciles targets from different administrative sources at district, state, and national levels.

The configurable sparsity penalty produces datasets of different sizes for different use cases. A high penalty ($\lambda_{L_0} = 10^{-4}$) retains approximately 50,000 records, suitable for national-level web-based simulation where download size and computation time matter. A low penalty ($\lambda_{L_0} = 10^{-8}$) retains approximately 3--4 million records, preserving geographic resolution for all 436 congressional districts.

The method is implemented in open-source Python packages and deployed as part of \policyengine{}'s US microsimulation model (\url{https://policyengine.org/us/model}), which uses these calibrated datasets for both national and subnational policy analysis.

The remainder of this paper is organized as follows. Section~\ref{sec:background} reviews survey calibration methods, spatial microsimulation, and the Hard Concrete distribution. Section~\ref{sec:data} describes the input microdata and calibration target sources. Section~\ref{sec:methodology} details the four-stage pipeline: clone creation, calibration matrix construction, $L_0$ optimization, and dataset assembly. Section~\ref{sec:results} presents calibration accuracy, sparsity analysis, and comparisons with IPF and GREG. Section~\ref{sec:discussion} discusses trade-offs, limitations, and generalizability. Section~\ref{sec:conclusion} concludes.
